Chapter 5
---------

The code below illustrates how increasing the half-saturation density (Ro) of the Rozenburg-MacArthur consumer-resource model alters the stability of the system.

```{r eval=TRUE, echo=FALSE}
### Get everything ready to go
setwd("~/Dropbox/FoodWebs") # set working directory
library("deSolve") # load required library
library("seqinr")
source('~/Dropbox/FoodWebs/Models.R') # source in R-M C-R model
```

```{r eval=TRUE}
###########  set parameters and state variables for the R-M C-R model
# state variable values (initial values at beginning of "experiments")
R <- 0.6
C <- 0.1
i.state <- c(R=0.6,C=0.1) 

# parameter values
r <- 1.0 # per capita rate of increase in resource
K <- 2.0 # indicated as 1.0 in Fig 2.1 description, but I think this was a typo in the book.
e <- 0.5 # predator conversion efficiency
Ro <- 0.5 # half-saturation density of predator functional response
m <- 0.5 # mortality rate of predator
a <- 1.7 # initial attack rate of 1.3 instead of 1.2 more closely replicates the figures in the book
p.rm1 <- c(r = r, e = e, a = a, K = K, Ro = Ro, m=m) # create a vector for the parameters for experiment #1
```

The figures below are the results of three separate simulations, where the half-saturation density was decreased from 0.5 to 0.2 and all other parameters were held constant.
```{r eval=TRUE, echo=FALSE, fig.width=10, warning=FALSE}
#### setup experiment. This experiment essentially solves the model at the initial C and R densities, and takes the new C and R densities and reruns the model, and so on until the end of time.
Time <- 300 # set time scale
rm1 <- ode(i.state,1:Time, rmcr, p.rm1) # run the experiment 

# plot graphs side by side
par(mfrow=c(1,3)) # plot graphs side-by-side

#### create resource isocline
Rx <- seq(0.1,2,0.1) # manipulating different Resource densities to solve R isocline.
Riso <- expression(r/a * (Rx + Ro) * (1 - Rx/K)) # set R = 0 and solved algebraically
RisoStable <- eval(Riso)

### create consumer isocline
Ciso <- expression(m * Ro / (e * a - m)) # set C = 0, and solved algebraically.
CisoStable <- eval(Ciso)

## Experiment 1: Ro = 0.5
Ro <- 0.5
p.rm1 <- c(r = r, e = e, a = a, K = K, Ro = Ro, m=m)
# plot consumer functional response and consumer isocline
Cfr <- expression(a * C * Rx / (Rx + Ro))
CfrData <- eval(Cfr)
plot(Rx,CfrData, ylim=c(0,0.18), type = "l", ylab="Consumption Rate of Consumer", xlab="Resource Density")
abline(v=CisoStable, lty = 2, col =2)
rect(CisoStable,0,max(Rx),0.18, col=col2alpha("red",0.5)) # shaded area indicates where consumer population growth rate is greater than zero

# plot stability around consumer and resource isoclines
plot(Rx,RisoStable, type = "l", ylab = "C", ylim=0:1, xlim = c(0,2), xlab="R", main="Half-saturation density = 0.5")
abline(v=CisoStable, lty = 2, col =2) 
legend("topleft", c("R-isocline","C-isocline"), lty=1:2, bty="n", cex=0.8, col=1:2)
points(i.state[1],i.state[2]) # starting point of experiment
arrows(rm1[-Time,2], rm1[-Time,3], rm1[-1,2], rm1[-1,3], length=0.1, lty=1) # trace stability across different time steps. Don't know why there are so many warnings

# plot densities
matplot(rm1[,1],rm1[,c(2,3)], type = "l", ylab="Density", xlab="Time", ylim=c(0,2))
legend("right", c("R","C"), lty=1:2, col=1:2, bty="n")

######## Experiment 2: Ro = 0.3
###  adjust parameters
Ro <- 0.3
p.rm2 <- c(r = r, e = e, a = a, K = K, Ro = Ro)
rm2 <- ode(i.state,1:Time, rmcr, p.rm2)

### adjust resource and consumer isoclines. IDEAS TO MAKE THIS LESS REDUNDANT???
Riso <- expression(r/a * (Rx + Ro) * (1 - Rx/K)) # adjusted attack rate
RisoStable <- eval(Riso)
Ciso <- expression(m * Ro / (e * a - m)) # adjusted attack rate
CisoStable <- eval(Ciso)

# plot consumer functional response and consumer isocline
Cfr <- expression(a * C * Rx / (Rx + Ro))
CfrData <- eval(Cfr)
plot(Rx,CfrData, ylim=c(0,0.18), type = "l", ylab="Consumption Rate of Consumer", xlab="Resource Density")
abline(v=CisoStable, lty = 2, col =2)
rect(CisoStable,0,max(Rx),0.18, col=col2alpha("red",0.5)) # shaded area indicates where consumer population growth rate is greater than zero

#### plot stabilities around consumer and resource isoclines.
plot(Rx, RisoStable, type = "l", ylab = "C", ylim=0:1, xlab="R", main="Half-saturation density = 0.3")
abline(v=CisoStable, lty = 2, col =2) 
legend("topright", c("R-isocline","C-isocline"), lty=1:2, bty="n", cex=0.8, col=1:2)
points(i.state[1],i.state[2]) # starting point of experiment
arrows(rm2[-Time,2], rm2[-Time,3], rm2[-1,2], rm2[-1,3], length=0.1, lty=1)

#### plot consumer and resource densities
matplot(rm2[,1],rm2[,c(2,3)], type = "l", ylab="Density", xlab="Time", ylim=c(0,2))
legend("topright", c("R","C"), lty=1:2, col=1:2, bty="n")



######## Experiment 3: Ro = 0.2
### adjust parameter values
Ro <- 0.2
p.rm3 <- c(r = r, e = e, a = a, K = K, Ro = Ro)
Time <- 100 # same pattern at larger time values, but I adjusted it to see the pattern more clearly
rm3 <- ode(i.state,1:Time, rmcr, p.rm3)

### adjust consumer and resource isocline
Riso <- expression(r/a * (Rx + Ro) * (1 - Rx/K)) 
RisoStable <- eval(Riso)
Ciso <- expression(m * Ro / (e * a - m)) 
CisoStable <- eval(Ciso)

# plot consumer functional response and consumer isocline
Cfr <- expression(a * C * Rx / (Rx + Ro))
CfrData <- eval(Cfr)
plot(Rx,CfrData, ylim=c(0,0.18), type = "l", ylab="Consumption Rate of Consumer", xlab="Resource Density")
abline(v=CisoStable, lty = 2, col =2)
rect(CisoStable,0,max(Rx),0.18, col=col2alpha("red",0.5)) # shaded area indicates where consumer population growth rate is greater than zero

### plot stabilities around resouce and consumer isoclines 
plot(Rx,RisoStable, type = "l", ylab = "C", ylim=0:1, xlab="R",main="Half-saturation Density = 0.2")
abline(v=CisoStable, lty = 2, col = 2) # attack rate adjusted to a = 2.0
legend("topright", c("R-isocline","C-isocline"), lty=1:2, col=1:2, bty="n", cex=0.8)
points(i.state[1],i.state[2]) # starting point of experiment
arrows(rm3[-Time,2], rm3[-Time,3], rm3[-1,2], rm3[-1,3], length=0.1, lty=1)

### plot densities
matplot(rm3[,1],rm3[,c(2,3)], type = "l", ylab="Density", xlab="Time", ylim=c(0,2))
legend("topright", c("R","C"), lty=1:2, col=1:2, bty="n", cex=0.8)
```

Below is a bifurcation plot, showing how the dynamics of the system changes along a continuous gradient of half-saturation densities.  Note that at high half-saturation densities, the consumers go extinct, but the stability of the system quickly changes over a short parameter range.
```{r eval=TRUE, echo=FALSE}
par(mfrow=c(1,2))
############  Figure 2.7a: Bifurcation plot  (All credit for replicating this goes to this website: http://www.r-bloggers.com/r-tools-for-dynamical-systems-bifurcation-plot-in-r%C2%A0for%C2%A0system%C2%A0of%C2%A0odes/)

param.name <- "Ro" # choose parameter to perturb
param.seq <- seq(1,0.1,length=50) # choose range of parameters
Ro <- 1.0
p.rm4 <- c(r = r, e = e, a = a, K = K, Ro = Ro, m=m) # set starting parameters.
param.index <- which(param.name == names(p.rm4)) # tells the loop which parameter in "p.rm4" to grab for manipulation.
Time <- 100 # decreased time to see outcomes of simulations

out <- list()
for (i in 1:length(param.seq))
  out[[i]] <- matrix(0, Time, length(i.state))

for (i in 1:length(param.seq)) {
  # set params
  p.rm4.loop <- p.rm4
  p.rm4.loop[param.index] <- param.seq[i] # changes the parameter value for manipulation in the "init" function below.
  # converge
  init <- ode(i.state, 1:Time, rmcr, p.rm4.loop)
  # get converged points
  out[[i]] <- ode(init[Time,-1], 1:Time, rmcr, p.rm4.loop)[,-1]
}

range.lim <- lapply(out, function(x) apply(x, 2, range)) # don't completely understand what is going on here...
range.lim <- apply(do.call("rbind", range.lim), 2, range) # this apparently doesn't work, but it doesn't mess things up... 

####  consumer bifurcation plot
plot.variable <- "C" # choose which variable to show
plot(0, 0, pch = "", xlab = param.name, ylab = plot.variable, xlim = range(param.seq), ylim = c(0,1), main = "Bifurcation plot") # range.lim[,plot.variable]
for (i in 1:length(param.seq)) {
  points(rep(param.seq[i], Time), out[[i]][,plot.variable])
}

####  resource bifurcation plot
plot.variable <- "R" # choose which variable to show
plot(0, 0, pch = "", xlab = param.name, ylab = plot.variable, xlim = range(param.seq), ylim = c(0,2), main = "Bifurcation plot") # ylim = range.lim[,plot.variable]
for (i in 1:length(param.seq)) {
  points(rep(param.seq[i], Time), out[[i]][,plot.variable])
}
```

Jacobian Matrix analysis

```{r}
# parameter values
jacobian.roz.mac <- function(r = 1.0, K = 2.0, e = 0.5, Ro = 0.5, m = 0.5, a = 1.3){
#r <- 1.0 # per capita rate of increase in resource 
#K <- 2.0 # indicated as 1.0 in Fig 2.1 description,   but I think this was a typo in the book.
#e <- 0.5 # predator conversion efficiency  
#Ro <- 0.5 # half-saturation density of predator functional response
#m <- 0.5 # mortality rate of predator
#a <- 1.3 # initial attack rate of 1.3 instead of 1.2 more closely replicates the figures in the book

  dR.dt <- expression(r * R * (1 - R / K) - a * C * R / (R + Ro))
  dC.dt <- expression(e * a * C * R / (R + Ro) - m * C)
  
  CRjac1 <- list(D(dR.dt,"R"), D(dC.dt, "R"), D(dR.dt, "C"), D(dC.dt, "C")) # making a list of the partial derivatives for the consumer resource model
  R <- m * Ro / (e * a - m) # this doesn't make sense...because it is the consumer isocline
  C <- eval(Riso)
    
  matrix(sapply(CRjac1, function(partial.deriv) eval(partial.deriv)), nrow = 2)
  }


jacobian.roz.mac(a=3)
seq.attack <- seq(1.1,3,by=0.1)
jacob.attackrate.change <- lapply(seq.attack,function(as) jacobian.roz.mac(a=as))
attack.rate.effect.res <- sapply(jacob.attackrate.change,"[",i=1,j=1)
plot(attack.rate.effect.res,type='l')
res.positive.attack.rate <- seq.attack[min(which(attack.rate.effect.res>0))]
max.real.eigen <- sapply(jacob.attackrate.change,function(x) max(Re(eigen(x)[["values"]])))

plot(max.real.eigen~seq.attack,type='l')
abline(h=0)
abline(v=1.7,lty=2)
```