---
title: "oscillation_metric"
author: "Matthew Barbour"
date: "February 24, 2015"
output: word_document
---

Load required R libraries for running the models.
```{r load libraries}
require(deSolve)
require(seqinr)
```

Setup the Rosenzweig-MacArthur Consumer-Resource model as formulated in Johnson and Amarasekare 2015. 
```{r}
## Rosenzweig-MacArthur consumer-resource model 
# assumes logistic resource growth and a type 2 functional response by the consumer (i.e. consumption rate saturates with resource density). This function is necessary for the ordinary differential equation solver, ode() 
rmcr <- function(t,y,p) {
  # t,y,p is the necessary format for solving using the ode() function in R
  R <- y[1]
  C <- y[2]
  with(as.list(p), {
    dR.dt <- r * R * (1 - q * R) - a * C * R / (1 + a * C * h * R)
    dC.dt <- e * a * C * R / (1 + a * C * h * R) - d * C
    return(list(c(dR.dt,dC.dt)))
  })
}
```

Set intial state variable (Resource and Consumer densities) for running the model. Note that it is not immediately clear what their initial state variable were so I just played around until I got ones that were close.
```{r}
# state variable values (initial values at beginning of "experiments")
R <- 1
C <- 0.2
i.state <- c(R = R, C = C) 
```

Replicate Fig. 3a
```{r}
## Experiment 1: a * h / q < 1

# parameter values for Harlequin bug and Trissolocus interaction, taken from Fig. 3 in Johnson & Amarasekare 2015
r <- 0.85 # per capita rate of increase in resource
q <- 1/16 # estimate competition coefficient from field data for Harlequin bug
e <- 1 # predator conversion efficiency
h <- 0.015 # handling time
d <- 0.33 # mortality rate of predator
a <- 1.11 # initial attack rate of 1.3 instead of 1.2 more closely replicates the figures in the book

a * h / q # 0.2664, which is lower than what was reported in the manuscript (0.35)

# run the experiment. This experiment essentially solves the model at the initial C and R densities, and takes the new C and R densities and reruns the model, and so on until the end of time.
p.rm1 <- c(r = r, e = e, a = a, q = q, h = h, d = d) # create a vector for the parameters for experiment #1
 
Time <- 1000 # set time scale
rm1 <- ode(i.state,1:Time, rmcr, p.rm1) # run the experiment 

# replicate Fig. 3a
matplot(rm1[,1], rm1[,c(2,3)],
        type = "l", ylab="Density", xlab="Time", ylim=c(0, C*15), main = "Trissolcus interaction")
legend("top", c("R","C"), lty=1:2, col=1:2, bty="n")
```

Replicate Fig. 3b. 
```{r }
##  parameters for Ooencyrtus interaction
h <- 0.05 
a <- 1.32 
e <- 1.82 
d <- 0.25 

a * h / q # now is above threshold (i.e. > 1)

# set parameters and run the model for the same amount of time.
p.rm2 <- c(r = r, e = e, a = a, h = h, q = q)
rm2 <- ode(i.state, 1:Time, rmcr, p.rm2)

# plot densities
matplot(rm2[,1],rm2[,c(2,3)], 
        type = "l", ylab="Density", xlab="Time", ylim=c(0,C*15), main = "Ooencyrtus interaction")
legend("top", c("R","C"), lty=1:2, col=1:2, bty="n")
```

After replicating these figures, it isn't exactly clear to me why a*h/q < 1 is a threshold. It seems like both interactions are oscillating...