---
title: "rosenzweig_macarthur_apparent_competition"
author: "Matthew Barbour"
date: "February 26, 2015"
output: word_document
---

Create model

Load required R libraries for running the models.
```{r load libraries}
require(deSolve)
require(seqinr)

source('~/Dropbox/FoodWebs/consumer_resource_models.R')
```

Set intial state variable (Resource and Consumer densities) for running the model. Note that it is not immediately clear what their initial state variable were so I just played around until I got ones that were close.

Apparent competition model
```{r}
# state variable values (initial values at beginning of "experiments")
R1 <- R2 <- 1
C1 <- 0.2
i.state <- c(R1 = R1, R2 = R2, C1 = C1) 

# parameter values
r1 <- r2 <- 1 # per capita rate of increase in resource
K1 <- K2 <- 2 # carrying capacity
e1 <- 0.3 # predator conversion efficiency
h1 <- h2 <- 0.03 # handling time
dC1 <- 0.5 # mortality rate of predator
a1 <- a2 <- 1 # attack rate of Consumer

# run the experiment. This experiment essentially solves the model at the initial C and R densities, and takes the new C and R densities and reruns the model, and so on until the end of time.
p.rm1 <- c(r1 = r1, r2 = r2, 
           K1 = K1, K2 = K2,
           e1 = e1, 
           a1 = a1*3, a2 = a2*3,
           h1 = h1*10, h2 = h2*10, 
           dC1 = dC1) # create a vector for the parameters for experiment #1
 
Time <- 400 # set time scale
rm1 <- ode(i.state, 1:Time, RM_apparent, p.rm1) # run the experiment 

# replicate Fig. 3a
matplot(rm1[,1], rm1[,c(2,3,4)],
        type = "l", ylab="Density", xlab="Time", ylim=c(0, C*15), main = "Apparent Competition", lwd = 3)
legend("top", c("R1","R2", "C"), lty=1:3, col=1:3, bty="n", lwd = 3)
```

2 consumer 2 resource model
```{r}
# new state variables for model
C2 <- C1
i.state2 <- c(R1 = R1, R2 = R2, C1 = C1, C2 = C2) 

# new parameters for model
e2 <- e1
dC2 <- dC1


# run the experiment. This experiment essentially solves the model at the initial C and R densities, and takes the new C and R densities and reruns the model, and so on until the end of time.
p.rm2 <- c(r1 = r1, r2 = r2, 
           K1 = K1, K2 = K2,
           e1 = e1, e2 = e2,
           a1 = a1*3, a2 = a2,
           h1 = h1, h2 = h2, 
           dC1 = dC1, dC1 = dC2) # create a vector for the parameters for experiment #2
 
rm2 <- ode(i.state2, 1:Time, RM_2C_2R, p.rm2) # run the experiment 

# replicate Fig. 3a
matplot(rm2[,1], rm2[,c(2,3,4,5)],
        type = "l", ylab="Density", xlab="Time", ylim=c(0, C*15), main = "2 Consumers, 2 Resources", lwd = 3)
legend("top", c("R1","R2", "C1", "C2"), lty=1:4, col=1:4, bty="n", lwd = 3)
```

This 2 consumer, 2 resource model appears quite robust to increases in attack rates or increases in handling time. In other words, it takes quite a lot to make this system oscillate, same with the apparent competition module. 

I wonder though if it is more structurally stable than the apparent competition module? I think this it can handle asymmetry in attack rates between predators (i.e., it can handle specialization). But perhaps I'm cheating here. I may need to add all four attack rates to see if this is true or not.

I wonder if I can determine the oscillation criteria for these 2 different modules?

Adding an additional predator though doesn't not change resource density (I'm assuming because exploitative competition amongs predators just reduces their respective densities).

2 consumer 2 resource FULL model. Doesn't seem to be working for sure right now...
```{r}
# new state variables for model
R1 <- 1
R2 <- 1
C1 <- 0.2
C2 <- 0.2
i.state3 <- c(R1 = R1, R2 = R2, C1 = C1, C2 = C2) 

# new parameters for model
r1 <- r2 <- 1
K1 <- K2 <- 2
e11 <- e12 <- e21 <- e22 <- 0.5
a11 <- a22 <- 1
a12 <- a21 <- 1
h11 <- h12 <- h21 <- h22 <- 0.5
dC1 <- dC2 <- 0.5 


# run the experiment. This experiment essentially solves the model at the initial C and R densities, and takes the new C and R densities and reruns the model, and so on until the end of time.
p.rm3 <- c(r1 = r1, r2 = r2, 
           K1 = K1, K2 = K2,
           e11 = e11, e22 = e22, e12 = e12, e21 = e21,
           a11 = a11, a22 = a22, a12 = a12, a21 = a21,
           h11 = h11, h22 = h22, h12 = h12, h21 = h21, 
           dC1 = dC1, dC1 = dC2)
 
Time <- 200
rm3 <- ode(i.state3, 1:Time, RM_2C_2R_full, p.rm3) # run the experiment 

# replicate Fig. 3a
matplot(rm3[,1], rm3[,c(2,3,4,5)],
        type = "l", ylab="Density", xlab="Time", ylim=c(0, 2), main = "2 Consumers, 2 Resources", lwd = 3)
legend("top", c("R1","R2", "C1", "C2"), lty=1:4, col=1:4, bty="n", lwd = 3)
```

Odd, increasing the handling time (for all) actually appears to stabilize the interaction.

Need to perform different experiments. 
- Character displacement
- Character Convergence

Need to read Abrams 1986, he analyzes linear functional responses, but the model is more comprehensive in other respects. Need to decide what is the best to do.